---
title: "R CCC analysis"
output: html_notebook
---
This is the CCC analysis notebook, divided in two main parts *Brain* and *Lung*
It starts by Brain, loading the data, cleaning and merging the data, transforming the corresponding columns for the analysis and finding the best comparable scales among manual score [0-4] to digital scores (%), where in Brain we have a max of 20% while in Lung we have a max of 60%. Therefore, using a comparable scale to meet the assumptions for the CCC analysis, we will establish.
For *Brain* [0-4] corresponding to (-Inf,0,5,10,15,20,Inf)(%), while for *Lung* [0-4] we will establish to (-Inf,15,30,45,60,Inf)(%)
# Load libraries


```{r}
gc()
rm(list = ls(all.names = TRUE))
output_dir <- "C:/Users/emari/OneDrive - Universitaet Bern/Nedim/2026/Results/CCC/"
if (!dir.exists(output_dir)){dir.create(output_dir,recursive = TRUE)}
```


```{r}
.libPaths("C:/Users/emari/OneDrive - Universitaet Bern/Documents/R/4.4.0/")
#BiocManager::install(c("DescTools","infotheo"))
library("ggplot2")
library("ggpubr")
library("dplyr")
library("tidyr")
library("DescTools")
library("infotheo")
```




## Reload brain data 
```{r}
sample_sheet = "C:/Users/emari/OneDrive - Universitaet Bern/Nedim/2026/data/All lungs and brain_cleaned_Dec2025_NK.xlsx"
sample_sheet_sheet_4_brain = readxl::read_excel(sample_sheet, col_names = TRUE, sheet = 4)
head(sample_sheet_sheet_4_brain)
colnames(sample_sheet_sheet_4_brain)


sample_sheet_sheet_6_brain = readxl::read_excel(sample_sheet, col_names = TRUE, sheet = 6)
head(sample_sheet_sheet_6_brain)
colnames(sample_sheet_sheet_6_brain)

sample_sheet_sheet_4_brain_coi <- sample_sheet_sheet_4_brain  |> dplyr::select(ID,`ID IVI`,`% Positive area`)
colnames(sample_sheet_sheet_4_brain_coi)
colnames(sample_sheet_sheet_6_brain)

# Need to merge colun % Positive Area from sheet 4 with Brain to plot sheet 6
sample_sheet_sheet_6_brain <- sample_sheet_sheet_6_brain |> dplyr::rename("ID"="ID...39")
sample_sheet_brain <- sample_sheet_sheet_6_brain |> dplyr::left_join(sample_sheet_sheet_4_brain_coi ,by = "ID")#c("ID...1"="ID"))

colnames(sample_sheet_brain)
head(sample_sheet_brain$`% Positive area.y`)


table(sample_sheet_brain$`Vaccinated (v), none (n)`)
table(sample_sheet_brain$`ID IVI.x`,sample_sheet_brain$`Euthanasia day (dpi)`)
unique(sample_sheet_brain$`Euthanasia day (dpi)`)
sample_sheet_brain <- sample_sheet_brain %>% dplyr::filter(!`Euthanasia day (dpi)` %in% c("7 p.v","6 p.v"))


sample_sheet_brain$Euthanasia_day_dpi = gsub(x = sample_sheet_brain$`Euthanasia day (dpi)`,
                                             pattern ="?p\\.v$",
                                             replacement = "")
sample_sheet_brain$Euthanasia_day_dpi <- as.numeric(sample_sheet_brain$Euthanasia_day_dpi)
  

message("using latest day of euthanasia day")
multima_day <- max(sample_sheet_brain$Euthanasia_day_dpi)

sample_sheet_brain <- sample_sheet_brain  |> dplyr::mutate(
    # Time group based on euthanasia day
    Time_Group = dplyr::case_when(
      Euthanasia_day_dpi >= 0 & Euthanasia_day_dpi <= 4 ~ "Early infection (1-4 dpi)",
      Euthanasia_day_dpi >= 5 & Euthanasia_day_dpi <= 7 ~ "Peak of disease (5-7 dpi)",
      Euthanasia_day_dpi >= 8 & Euthanasia_day_dpi <= multima_day  ~ paste0("Resolution (8-",multima_day," dpi)"),
      TRUE ~ "Other"),
    # Vaccination status
    Vaccination_Status = dplyr::if_else(`Vaccinated (v), none (n)` == "n", "n", "v") |> factor())#(n)`, levels = c("n", "v"))
  
sample_sheet_brain <- sample_sheet_brain |> dplyr::mutate( Vaccination_Status = dplyr::case_when( grepl(" p\\.v$", `Euthanasia day (dpi)`) ~ "v/nc", TRUE ~ as.character(Vaccination_Status) ),
                                                     Vaccination_Status = factor(Vaccination_Status, levels = c("n", "v", "v/nc")))

table(sample_sheet_brain$Vaccination_Status)
# Create combined grouping variable
sample_sheet_brain <- sample_sheet_brain  |> dplyr::mutate(Combined_Group = paste0(Time_Group, "-", Vaccination_Status))

sample_sheet_brain$sex <- as.factor(toupper(sample_sheet_brain$sex))


# Prepare data for comparison
comparison_data_brain <- sample_sheet_brain |> 
  dplyr::select(
    `Manual Score brain`,
    ID...1,
    sex,
    `qPCR swabs (RNA/ml)`,
    `Histopathological score brain`,
    #`Histopathological score brain...24`,
    `% Positive area.y`,
    Time_Group,
    Vaccination_Status,
    `qPCR Brain (RNA/ml)`,
    #qPCR brain log10,
    `Live Virus log10 brain`,
    `Live virus (PFU/mouse) brain`) |>
  tidyr::drop_na(`% Positive area.y`)



# Rename for easier plotting
names(comparison_data_brain) <- c(
  "manual semiquantification brain",
  "ID",
  "sex",
  "qPCR_swabs",
  "Histopath_Brain",
  #"Histopath_Brain_24",
  "digital quantification brain",
  "Time_Group",
  "Vaccination_Status",
  "qPCR Brain (RNA/ml)",
  #qPCR brain log10,
  "Live Virus log10 brain",
  "Live virus (PFU/mouse) brain")


#change scores from 0-5 to 0-4 (leave the only animal with score 1 as it is and change the rest.)
comparison_data_brain$`manual semiquantification brain`

comparison_data_brain$`manual semiquantification brain` <- dplyr::recode(
  comparison_data_brain$`manual semiquantification brain`,
  `0` = 0,
  `2` = 1,
  `3` = 2,
  `4` = 3,
  `5` = 4
)

comparison_data_brain$`manual semiquantification brain`
#Treat v/nc as vaccinated --> converted
comparison_data_brain$Vaccination_Status
comparison_data_brain$Vaccination_Status <- if_else(comparison_data_brain$Vaccination_Status == "v/nc","v",comparison_data_brain$Vaccination_Status)

comparison_data_brain
```

```{r}
sample_sheet_brain
```


```{r}
sample_sheet_lungs = readxl::read_excel(sample_sheet, col_names = TRUE, sheet = 2)
head(sample_sheet_lungs)
colnames(sample_sheet_lungs)
```
```{r}
head(sample_sheet_brain)
colnames(sample_sheet_lungs)
```


```{r}
sample_sheet_common_brain_and_lung <- sample_sheet_brain %>% dplyr::inner_join(sample_sheet_lungs,by = c("ID...1"="ID"))
sample_sheet_common_brain_and_lung
output_dir = "C:/Users/emari/OneDrive - Universitaet Bern/Nedim/2026/data/"

write.table(file=paste0(output_dir,"common_brain_and_lung.csv"),
            x=sample_sheet_common_brain_and_lung,
            sep=",")

```



```{r}
# Correct data from brain not log-ed.
message("Some values in the qPCR for swabs and Brain (RNA/ml) and Live virus, are converted to log10, therefore ok, while others have a vlaue that is huge and need to be converted.")
head(comparison_data_brain$qPCR_swabs)
head(comparison_data_brain$`qPCR Brain (RNA/ml)`)
head(comparison_data_brain$`Live virus (PFU/mouse) brain`)

```
```{r}
message("Tipycally values smaller than 100 are log10 converted...")
log_candidates <- c("qPCR_swabs","qPCR Brain (RNA/ml)", "Live virus (PFU/mouse) brain")
for (col in log_candidates) {
  comparison_data_brain[[col]] <- ifelse(
    comparison_data_brain[[col]] > 100,
    log10(comparison_data_brain[[col]]),
    comparison_data_brain[[col]])}
```

```{r}
head(comparison_data_brain$qPCR_swabs)
min(round(comparison_data_brain$qPCR_swabs[!is.na(comparison_data_brain$qPCR_swabs)],digits = 3))
max(round(comparison_data_brain$qPCR_swabs[!is.na(comparison_data_brain$qPCR_swabs)],digits = 3))
min(round(comparison_data_brain$qPCR_swabs,digits = 3))
head(comparison_data_brain$`qPCR Brain (RNA/ml)`)
min(round(comparison_data_brain$`qPCR Brain (RNA/ml)`[!is.na(comparison_data_brain$`qPCR Brain (RNA/ml)`)],digits = 5))
max(round(comparison_data_brain$`qPCR Brain (RNA/ml)`[!is.na(comparison_data_brain$`qPCR Brain (RNA/ml)`)],digits = 5))
#comparison_data$`qPCR Brain (RNA/ml)`[round(comparison_data$`qPCR Brain (RNA/ml)`[!is.na(comparison_data$`qPCR Brain (RNA/ml)`)],digits = 3) == 68.616]
#comparison_data$`qPCR Brain (RNA/ml)`[comparison_data$`qPCR Brain (RNA/ml)` == 68.616] found and converted to the real value from Nedim data.

head(comparison_data_brain$`Live virus (PFU/mouse) brain`)
```
```{r}
head(comparison_data_brain)
dim(comparison_data_brain)
comparison_data_brain
```


### Spearman
#### How variables change together? (meassured by correlation)

#### Brain manual correlation to digital
```{r}
brain_scores <- comparison_data_brain |>
  dplyr::select(
    manual = `manual semiquantification brain`,
    digital = `digital quantification brain`,
    vaccination_status = Vaccination_Status,
    time_group = Time_Group
  ) |>
  drop_na()

# Brain correlation in general manual to digital
cor_spearman_brain <- cor(brain_scores$manual, brain_scores$digital,method = "spearman", use = "complete.obs")
cor_spearman_brain


cor_test_brain <- cor.test(brain_scores$manual,
                           brain_scores$digital,
                           method = "spearman",
                           exact = FALSE)

paste0('Cor: ',round(cor_test_brain$estimate,5))      # Spearman rho
paste0('P-val: ',cor_test_brain$p.value,25)       # p‑value


```

```{r}
cor_brain_non_stratified <- ggplot(brain_scores, aes(x = manual, y = digital)) +
  geom_point(size = 3, alpha = 0.7, color = "darkblue") +
  geom_smooth(method = "lm", se = TRUE, color = "red", linewidth = 1) +
  ggpubr::stat_cor(
    method = "spearman",
    label.x.npc = "left",
    label.y.npc = "top",
    label.sep = "\n") +#,   # <-- ensures rho and p-value appear on separate lines
    #aes(label = paste0("Spearman = ", after_stat(r.label),"\nP = ", after_stat(p.label)))) +
  labs(title = "Manual vs Digital Brain Scores",
    x = "Manual semiquantification (brain)",
    y = "Digital quantification (brain)") +
  theme_bw(base_size = 10)
cor_brain_non_stratified
ggsave(paste0(output_dir,"cor_brain_non_stratified.pdf"),
       cor_brain_non_stratified,
       width = 6.5,
       height = 4.5)
```


```{r warning=FALSE}
# Brain correlatios scores manual to digital by time group and vaccination status.
brain_scores_cor_brain <- list()
for (vs in brain_scores$vaccination_status){
  for (tg in brain_scores$time_group){
    df <- brain_scores |> dplyr::filter(vaccination_status == vs & time_group == tg)
    brain_scores_cor_brain[[paste0(vs)]][[paste0(tg)]] <- cor(df$manual,df$digital,method = "spearman", use = "complete.obs")}}
brain_scores_cor_brain
```

```{r,fig.width=12,fig.height=8}
# Get number of samples per group
n_by_group <- brain_scores |>
      #dplyr::filter(!is.na(.data[[plt]]), .data[[plt]] != "") |>
      dplyr::group_by(time_group,vaccination_status) |>
      dplyr::summarise(n = dplyr::n(), .groups = "drop")
n_by_group

# Plot
ggplot(brain_scores, aes(x = manual, y = digital)) +
  #geom_point(aes(color = time_group), size = 3, alpha = 0.7) +
  geom_jitter(aes(color = time_group), size = 3, alpha = 0.7) +
  geom_smooth(aes(color = time_group), method = "lm", se = FALSE) +
  
  # Spearman correlation per facet
  ggpubr::stat_cor(
    method = "spearman",
    label.x.npc = "left",
    label.y.npc = "top",
    label.sep = "\n") +
    #aes(label = paste0("S = ", after_stat(r.label),"\nP = ", after_stat(p.label)))) +
  geom_text(data = n_by_group,aes(x = -Inf, y = Inf, label = paste0("n = ", n)),
            hjust = -0.1, vjust = 1.2, size = 3) +
  facet_grid(vaccination_status ~ time_group) +
  labs(title = "Manual vs Digital Brain Scores",
    x = "Manual semiquantification (brain)",
    y = "Digital quantification (brain)") +
  theme_bw(base_size = 10) +
  theme(legend.position = "none",
    strip.text = element_text(size = 10, face = "bold"))
```


# MI
#### Will tell you if the methods are related, but it cannot tell you if they are interchangeable. 
#### High MI just means the digital score is a good "predictor" of the manual score.

## Discrete
### Brain
```{r}
#install.packages("infotheo")
library(infotheo)

brain_disc <- discretize(brain_scores)
mi_value_brain_discrete <- mutinformation(brain_disc$manual, brain_disc$digital)
mi_value_brain_discrete
# need to run under before this..
# mi_value_brain_discrete_binned <- mutinformation(brain_disc$manual, brain_disc$digital_binned)
# mi_value_brain_discrete_binned
```


# CCC
### 1. Do the manual  and digital quantification agree, and can they be used interchangeably?  
### 2. Agreement measures how close the actual values are to one another


### 3.Convert the "%" scores from the digital to binary format.

```{r}
# Convert the digital to bins using 
library("rpart")
#BiocManager::install(c("rpart","rpart.plot"))
library("rpart.plot")

#rpart() fits a CART model (Classification And Regression Trees).
# If digital < 0.056 → predict manual = 0
# Else if digital < 0.468 → predict manual = 2
# Else if digital < 4.569 → predict manual = 3
# Else → predict manual = 4

table(brain_scores$manual)
fit <- rpart(manual ~ digital, data = brain_scores)
fit
cutpoints <- sort(unique(fit$splits[, "index"]))
cutpoints


# Now change the only one sample with vlaue 1 to 0.
brain_scores$manual_collapsed <- ifelse(brain_scores$manual == 1, 0, brain_scores$manual)
brain_scores$manual_collapsed <- factor(brain_scores$manual_collapsed)
table(brain_scores$manual_collapsed)
fit2 <- rpart(manual_collapsed ~ digital, data = brain_scores)
fit2
#rpart.plot::rpart.plot(fit2)
#rpart.plot(fit2)

pdf(paste0(output_dir,"brain_digital_score_cutpoints_rtree.pdf"))
rpart.plot(fit2)
dev.off()

plotcp(fit)
#fit2$splits
predicted_digital_scores <- predict(fit2, newdata = data.frame(digital=brain_scores$digital), type = "class")
cutpoints_df <- data.frame(binned_digital_scores_rtree=predicted_digital_scores,true_label_digital_score_brain=brain_scores$digital)
write.table(cutpoints_df,paste0(output_dir,"brain_cutpoints_digital_score_rtree.csv"),sep = ",")
cutpoints2 <- sort(unique(fit2$splits[, "index"]))
cutpoints2
brain_scores$digital_tree_binned <- cut(brain_scores$digital,
                                        breaks = c(-Inf, cutpoints2, Inf),
                                        labels = c(0,2,3,4),
                                        include.lowest = TRUE)

table(manual = brain_scores$manual_collapsed,
      digital = brain_scores$digital_tree_binned)
```
## kappa test, not really used..
```{r}
#BiocManager::install("psych")
library("psych")
psych::cohen.kappa(cbind(brain_scores$manual_collapsed, brain_scores$digital_tree_binned))#,weights = "quadratic")


```


```{r}
# Bin the digital scores to match the 0-5 scale
# Adjust 'breaks' based on your clinical expertise/standard and avoid arbitrary binning or use as above rpart?
message("use this if we want to set the cutpoints of the scores manually")
# brain_scores$digital_binned <- cut(brain_scores$digital,
#                                    #breaks = c(-Inf, 1, 5, 10, 15, 20, Inf),
#                                    breaks = c(-Inf, 5, 10, 15, 20, Inf),
#                                    labels = 0:4,
#                                    include.lowest = TRUE)


message('run from here to use the rpart cuts done to the data, rather than using our "bins"')
# Ensure both are numeric or factors for the function
brain_scores$manual
brain_scores$manual_collapsed
#brain_scores$digital_binned <- as.numeric(as.character(brain_scores$digital_binned))
brain_scores$manual_collapsed <- as.numeric(as.character(brain_scores$manual_collapsed))
brain_scores$digital_tree_binned <- as.numeric(as.character(brain_scores$digital_tree_binned))
#max(brain_scores$digital_binned)
max(brain_scores$digital_tree_binned)
```


```{r warning=FALSE}
#install.packages("DescTools")
library("DescTools")
#--> Results :  CCC estimate,accuracy,precision,95% CI

#ccc_result_brain <- CCC(brain_scores$manual, brain_scores$digital)
#for status_vaccinated in brain_scores

ccc_brain_scores_stratified <- list()
for (vs in brain_scores$vaccination_status){
  for (tg in brain_scores$time_group){
    df <- brain_scores |> dplyr::filter(vaccination_status == vs & time_group == tg)
    
    ccc_score <- CCC(df$manual_collapsed,df$digital_tree_binned)
    ccc_brain_scores_stratified[[paste0(vs)]][[paste0(tg)]] <- round(ccc_score$rho.c[,1],3)}}

ccc_brain_scores_stratified
```
### plot brains ccc stratified
```{r}
# add CCC score to each group..
#ccc_stratified_scores <- ccc_brain_scores_stratified
ccc_brains_df <- do.call(rbind,
                  lapply(names(ccc_brain_scores_stratified), function(vs) {data.frame(
                    vaccination_status = vs,
                    time_group = names(ccc_brain_scores_stratified[[vs]]),
                    ccc_score = unlist(ccc_brain_scores_stratified[[vs]]),
                    row.names = NULL,
                    stringsAsFactors = FALSE)}))




#ccc_by_group <- ccc_brain_scores_stratified
brain_ccc_strat <- ggplot(brain_scores, aes(x = manual_collapsed, y = digital_tree_binned)) +
  #geom_jitter(size = 3, alpha = 0.7, color = "#3C1642") +
  geom_point(size = 3, alpha = 0.7, color = "#3C1642") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  stat_cor(method = "spearman", label.x =0, label.y = 5,size=5) +
  #annotate("text", label = paste0("CCC = ", round(ccc_result_brain$rho.c[,1], 3)), x= 0.44,y=3.7,#hjust=-0.3,vjust=3,  size=5) +
  geom_text(data = ccc_brains_df,aes(x = -Inf, y = Inf, label = paste0("CCC = ", ccc_score)),
            hjust = -0.1, vjust = 1.2, size = 3) +
  facet_grid(vaccination_status ~ time_group) +
  xlim(0,4)+
  ylim(0,4)+
  labs(
    title = "Manual vs Digital Brain Scoring",
    x = "Manual Semiquantification",
    y = "Digital Quantification (% Positive Area-Binnned)",
    subtitle = paste0("n=",brain_scores |> summarize(n=n())) #lung_scores |> count()
  ) +
  theme_bw()
brain_ccc_strat
```


```{r warning=FALSE}
#ccc_result_brain <- CCC(brain_scores$manual, brain_scores$digital_binned)
ccc_result_brain <- CCC(brain_scores$manual_collapsed, brain_scores$digital_tree_binned)
ccc_result_brain
round(ccc_result_brain$rho.c[,1],3)
ccc_result_brain$rho.c      # overall CCC
ccc_result_brain$s.shift    # scale shift
ccc_result_brain$l.shift    # location shift
ccc_result_brain$C.b        # accuracy
ccc_result_brain$rho.c$lwr.ci      # lower CI
ccc_result_brain$rho.c$upr.ci      # upper CI
```

```{r}
write.table(comparison_data_brain,file = paste0(output_dir,"comparison_data_brain.csv"),sep = ",")
```


```{r}
#brain_ccc <- ggplot(brain_scores, aes(x = manual, y = digital_binned)) +
brain_ccc <- ggplot(brain_scores, aes(x = manual_collapsed, y = digital_tree_binned)) +
  #geom_jitter(size = 3, alpha = 0.7, color = "#3C1642") +
  geom_point(size = 3, alpha = 0.7, color = "#3C1642") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  stat_cor(method = "spearman", label.x =0, label.y = 5,size=5) +
  annotate("text", 
           label = paste0("CCC = ", round(ccc_result_brain$rho.c[,1], 3),"\n95% CI:",round(ccc_result_brain$rho.c$lwr.ci,3),"-",round(ccc_result_brain$rho.c$upr.ci,3)),
           x= 0.44,y=3.7,#hjust=-0.3,vjust=3,
           size=5) +
  xlim(0,4)+
  ylim(0,4)+
  labs(
    title = "Manual vs Digital Brain Scoring",
    x = "Manual Semiquantification",
    y = "Digital Quantification (% Positive Area-Binnned)",
    subtitle = paste0("n=",brain_scores |> summarize(n=n())) #lung_scores |> count()
  ) +
  theme_bw()
brain_ccc
```


```{r}

```


```{r}
```


```{r}
```
## Reload data lung
```{r}
sample_sheet = "C:/Users/emari/OneDrive - Universitaet Bern/Nedim/2026/data/All lungs and brain_cleaned_Dec2025_NK.xlsx"
sample_sheet_lungs = readxl::read_excel(sample_sheet, col_names = TRUE, sheet = 2)
head(sample_sheet_lungs)
colnames(sample_sheet_lungs)
table(sample_sheet_lungs$`ID IVI`,sample_sheet_lungs$`Euthanasia day (dpi)`)

table(sample_sheet_lungs$`Vaccinated (v), none (n)`)
# sample_sheet_lungs$Euthanasia_day_dpi = gsub(x = sample_sheet_lungs$`Euthanasia day (dpi)`,
#                                              pattern ="?p\\.v$",
#                                              replacement = "")
sample_sheet_lungs$Euthanasia_day_dpi <- as.numeric(sample_sheet_lungs$`Euthanasia day (dpi)`)
mterm_val <- max(sample_sheet_lungs$Euthanasia_day_dpi)

#sample_sheet_lungs$`Vaccinated (v), none (n)`

message("Add to plot the number of samples such as: n=12")

sample_sheet_lungs <- sample_sheet_lungs  |> dplyr::mutate(
    # Time group based on euthanasia day
    Time_Group = case_when(
      Euthanasia_day_dpi >= 0 & Euthanasia_day_dpi <= 4 ~ "Early infection (1-4 dpi)",
      Euthanasia_day_dpi >= 5 & Euthanasia_day_dpi <= 7 ~ "Peak of disease (5-7 dpi)",
      Euthanasia_day_dpi >= 8 & Euthanasia_day_dpi <= mterm_val ~ paste0("Resolution (8-",mterm_val," dpi)"),
      TRUE ~ "Other"),
    # Vaccination status
    Vaccination_Status = if_else(`Vaccinated (v), none (n)` == "n", "n", "v") |> factor())

# Create combined grouping variable
sample_sheet_lungs <- sample_sheet_lungs  |> dplyr::mutate(
    Combined_Group = paste(Time_Group, "-", Vaccination_Status)
  )

table(sample_sheet_lungs$Vaccination_Status)

# Prepare data for comparison
comparison_data_lung <- sample_sheet_lungs |> 
  dplyr::select(
    `Manual Score (0 to 4)`,
    ID,
    `qPCR swabs (RNA/ml)`,
    `Histopathological score`,
    #`Histopathological score brain...24`,
    `% Positivity in total parenchyma`,
    Time_Group,
    Vaccination_Status,
    `qPCR Lung (RNA/ml)`,
    #qPCR brain log10,
    `Live virus (PFU/ml)`) #|>  drop_na(`% Positive area.y`)




# Rename for easier plotting
names(comparison_data_lung) <- c(
  "manual semiquantification lung",
  "ID",
  "qPCR_swabs",
  "Histopath_Lung",
  #"Histopath_Brain_24",
  "digital quantification lung",
  "Time_Group",
  "Vaccination_Status",
  "qPCR Lung (RNA/ml)",
  #qPCR brain log10,
  "Live virus (PFU/ml)")


head(comparison_data_lung)
dim(comparison_data_lung)
comparison_data_lung

```
# Spearman
### How variables change together? (meassured by correlation)
### Lung manual correlation to digital

```{r}
lung_scores <- comparison_data_lung |>
  dplyr::select(
    manual = `manual semiquantification lung`,
    digital = `digital quantification lung`,
    vaccination_status = Vaccination_Status,
    time_group = Time_Group)# |>drop_na()

cor_spearman_lung <- cor(lung_scores$manual, lung_scores$digital,
                    method = "spearman", use = "complete.obs")
cor_spearman_lung


cor_test_lung <- cor.test(lung_scores$manual,
                           lung_scores$digital,
                           method = "spearman",
                           exact = FALSE)

paste0('Cor: ',round(cor_test_lung$estimate,5))      # Spearman rho
paste0('P-val: ',cor_test_lung$p.value,25)       # p‑value

```

```{r}
cor_lung_non_stratified <- ggplot(lung_scores, aes(x = manual, y = digital)) +
  geom_jitter(size = 3, alpha = 0.7, color = "darkorange") +
  geom_smooth(method = "lm", se = TRUE, color = "red", linewidth = 1) +
  ggpubr::stat_cor(
    method = "spearman",
    label.x.npc = "left",
    label.y.npc = "top",
    labels.sep = "\n") +
    #aes(label = paste0("Spearman = ", after_stat(r.label),"\nP = ", after_stat(p.label)))) +
  labs(title = "Manual vs Digital Lung Scores",
    x = "Manual semiquantification (lung)",
    y = "Digital quantification (lung)") +
  theme_bw(base_size = 10)

cor_lung_non_stratified
ggsave(paste0(output_dir,"cor_lung_non_stratified.pdf"),
       cor_lung_non_stratified,
       width = 6.5,
       height = 4.5)
```

```{r}
# Get number of samples per group
n_by_group <- lung_scores |>
      #dplyr::filter(!is.na(.data[[plt]]), .data[[plt]] != "") |>
      dplyr::group_by(time_group,vaccination_status) |>
      dplyr::summarise(n = dplyr::n(), .groups = "drop")
n_by_group


ggplot(lung_scores, aes(x = manual, y = digital)) +
  geom_jitter(aes(color = time_group), size = 3, alpha = 0.7) +
  geom_smooth(aes(color = time_group), method = "lm", se = FALSE) +
  
  # Spearman correlation per facet
  ggpubr::stat_cor(
    method = "spearman",
    label.x.npc = "left",
    label.y.npc = "top",
    label.sep = "\n") +
    #aes(label = paste0("S = ", after_stat(r.label),"\nP = ", after_stat(p.label)))) +
  geom_text(data = n_by_group,aes(x = -Inf, y = Inf, label = paste0("n = ", n)),
            hjust = -0.1, vjust = 1.2, size = 3) +
  
  facet_grid(vaccination_status ~ time_group) +
  labs(title = "Manual vs Digital Lung Scores",
    x = "Manual semiquantification (lung)",
    y = "Digital quantification (lung)") +
  theme_bw(base_size = 10) +
  theme(legend.position = "none",
    strip.text = element_text(size = 10, face = "bold"))

```

```{r}
write.table(comparison_data_lung,file = paste0(output_dir,"comparison_data_lung.csv"),sep = ",")
```


```{r}
```



# MI
#### Will tell you if the methods are related, but it cannot tell you if they are interchangeable. 
#### High MI just means the digital score is a good "predictor" of the manual score.

## Discrete
### Lung
```{r}

lung_disc <- discretize(lung_scores)
mi_value_lung_discrete <- mutinformation(lung_disc$manual, lung_disc$digital)
mi_value_lung_discrete

```
## tree classifier for binning scores
```{r}
#rpart() fits a CART model (Classification And Regression Trees).
# If digital < 0.056 → predict manual = 0
# Else if digital < 0.468 → predict manual = 2
# Else if digital < 4.569 → predict manual = 3
# Else → predict manual = 4

table(lung_scores$manual)
fit <- rpart(manual ~ digital, data = lung_scores)
fit

cutpoints <- sort(unique(fit$splits[, "index"]))
cutpoints


lung_scores$manual <- factor(lung_scores$manual)
table(lung_scores$manual)
#rpart.plot::rpart.plot(fit2)
#rpart.plot(fit2)

pdf(paste0(output_dir,"lung_digital_score_cutpoints_rtree.pdf"))
rpart.plot(fit)
dev.off()


#fit2$splits
predicted_digital_scores <- predict(fit, newdata = data.frame(digital=lung_scores$digital), type = "class")
cutpoints_df <- data.frame(binned_digital_scores_rtree=predicted_digital_scores,true_label_digital_score_lung=lung_scores$digital)
write.table(cutpoints_df,paste0(output_dir,"lung_cutpoints_digital_score_rtree.csv"),sep = ",")


lung_scores$digital_tree_binned <- cut(lung_scores$digital,
                                        breaks = c(-Inf, cutpoints, Inf),
                                        labels = c(0:4),
                                        include.lowest = TRUE)

table(manual = lung_scores$manual,
      digital = lung_scores$digital_tree_binned)
```


```{r}
# lung_scores$digital_binned <- cut(lung_scores$digital,
#                                    breaks = c(-Inf, 10, 25, 40, 55, Inf),
#                                    labels = 0:4,
#                                    include.lowest = TRUE)


# Ensure both are numeric or factors for the function
lung_scores$manual
lung_scores$manual <- as.numeric(as.character(lung_scores$manual))
lung_scores$digital_tree_binned <- as.numeric(as.character(lung_scores$digital_tree_binned))
max(lung_scores$digital_tree_binned)
```
# CCC
### 1. Do the manual  and digital quantification agree, and can they be used interchangeably?  
### 2. Agreement measures how close the actual values are to one another


### 3.Convert the "%" scores from the digital to binary format.

```{r}
ccc_lung_scores_stratified <- list()
for (vs in lung_scores$vaccination_status){
  for (tg in lung_scores$time_group){
    df <- lung_scores |> dplyr::filter(vaccination_status == vs & time_group == tg)
    
    ccc_score <- CCC(df$manual,df$digital_tree_binned)
    ccc_lung_scores_stratified[[paste0(vs)]][[paste0(tg)]] <- round(ccc_score$rho.c[,1],3)}}

ccc_lung_scores_stratified
```
```{r}
#ccc_stratified_scores <- ccc_brain_scores_stratified
ccc_lungs_df <- do.call(rbind,
                  lapply(names(ccc_brain_scores_stratified), function(vs) {data.frame(
                    vaccination_status = vs,
                    time_group = names(ccc_brain_scores_stratified[[vs]]),
                    ccc_score = unlist(ccc_brain_scores_stratified[[vs]]),
                    row.names = NULL,
                    stringsAsFactors = FALSE)}))


lung_ccc_strat <- ggplot(lung_scores, aes(x = manual, y = digital_tree_binned)) +
  geom_jitter(size = 3, alpha = 0.7, color = "darkorange") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  stat_cor(method = "spearman", label.x =0, label.y = 5,size=5) +
  #annotate("text", label = paste0("CCC = ", round(ccc_result_brain$rho.c[,1], 3)), x= 0.44,y=3.7,#hjust=-0.3,vjust=3,  size=5) +
  geom_text(data = ccc_lungs_df,aes(x = -Inf, y = Inf, label = paste0("CCC = ", ccc_score)),
            hjust = -0.1, vjust = 1.2, size = 3) +
  facet_grid(vaccination_status ~ time_group) +
  xlim(0,4)+
  ylim(0,4)+
  labs(title = "Manual vs Digital Lung Scoring",
       x = "Manual Semiquantification",
       y = "Digital Quantification (% Positive Area-Binnned)",
       subtitle = paste0("n=",lung_scores |> summarize(n=n()))) + #lung_scores |> count()
  theme_bw()
lung_ccc_strat
```


```{r}
#install.packages("DescTools")
library(DescTools)
#--> Results :  CCC estimate,accuracy,precision,95% CI

#ccc_result_lung <- CCC(lung_scores$manual, lung_scores$digital)
#ccc_result_lung <- CCC(lung_scores$manual, lung_scores$digital_binned)
ccc_result_lung <- CCC(lung_scores$manual, lung_scores$digital_tree_binned)
ccc_result_lung
round(ccc_result_lung$rho.c[,1],3)
ccc_result_lung$rho.c      # overall CCC
ccc_result_lung$s.shift    # scale shift
ccc_result_lung$l.shift    # location shift
ccc_result_lung$C.b        # accuracy
ccc_result_lung$rho.c$lwr.ci #lower ci
ccc_result_lung$rho.c$upr.ci #upper ci
```
```{r}


#lung_ccc <- ggplot(lung_scores, aes(x = manual, y = digital_binned)) +
lung_ccc <- ggplot(lung_scores, aes(x = manual, y = digital_tree_binned)) +
  #geom_jitter(size = 3, alpha = 0.7, color = "#FB8500") +
  geom_point(size = 3, alpha = 0.7, color = "#FB8500") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  stat_cor(method = "spearman", label.x =0, label.y = 5,size=5) +
  annotate("text", 
           label = paste0("CCC = ", round(ccc_result_lung$rho.c[,1], 3),"\n95% CI:",round(ccc_result_lung$rho.c$lwr.ci,3),"-",round(ccc_result_lung$rho.c$upr.ci,3)),

           x=0.44, y=3.7,size=5) +
  xlim(0,4)+
  ylim(0,4)+
  labs(title = "Manual vs Digital Lung Scoring",
       x = "Manual Semiquantification",
       y = "Digital Quantification (% Positive Area-Binned)",
       subtitle = paste0("n=",lung_scores |> summarize(n=n()))) +  #lung_scores |> count()
  theme_bw()
  
lung_ccc
```


```{r,fig.width=11,fig.height=5}
# Combine into 2‑panel figure
two_panel <- ggarrange(lung_ccc,brain_ccc,
                       ncol = 2,
                       nrow = 1,
                       common.legend = FALSE)

two_panel

# Save the two panel plot
ggsave(paste0(output_dir, "brain_mq_vs_dq_&_lubg_mq_vs_dq_ccc.pdf"), 
       two_panel, 
       width = 13, 
       height = 4.5, 
       dpi = 600)
```

# Bland Altman Brain
```{r}


library(ggplot2)
manual  <- brain_scores$manual
digital <- brain_scores$digital_tree_binned

diffs <- manual - digital
means <- (manual + digital) / 2

bias <- mean(diffs)
loa  <- bias + c(-1.96, 1.96) * sd(diffs)

ba_plot_brain <- ggplot(data.frame(means, diffs), aes(x = means, y = diffs)) +
  #geom_point(size = 3, alpha = 0.7, color = "#219ebc") +
  geom_jitter(size = 3, alpha = 0.7, color = "#3C1642") +
  geom_hline(yintercept = bias, color = "red", size = 1) +
  geom_hline(yintercept = loa, linetype = "dashed", color = "black") +
  annotate("text", x = 0.5, y = bias + 0.2,
           label = paste0("Bias = ", round(bias, 2))) +
  annotate("text", x = 0.5, y = loa[1] + 0.2,
           label = paste0("Lower LoA = ", round(loa[1], 2))) +
  annotate("text", x = 0.5, y = loa[2] - 0.2,
           label = paste0("Upper LoA = ", round(loa[2], 2))) +
  labs(title = "Bland–Altman Plot: Manual vs Digital (Binned)",
       x = "Mean of Methods",
       y = "Difference (Manual - Digital)") +
  theme_bw()

ba_plot_brain
```


# Bland Altman Lung
```{r}
library(DescTools)

# Bland–Altman analysis
ba_lung <- data.frame("Manual"=lung_scores$manual,"Digital"=lung_scores$digital_tree_binned)
ba_lung$avg <- rowMeans(ba_lung)
ba_lung$diff <- ba_lung$Manual - ba_lung$Digital

mean_diff <- mean(ba_lung$diff) # Mean difference
lower <- mean_diff - 1.96 * sd(ba_lung$diff) # Lower limit
upper <- mean_diff + 1.96 * sd(ba_lung$diff) # Upper limit

library(ggplot2)
ggplot(ba_lung, aes(x = avg, y = diff)) +
 #geom_jitter(size = 2) +
 geom_jitter(size = 3, alpha = 0.7, color = "#FB8500") +
 geom_hline(yintercept = mean_diff, color = "blue") +
 geom_hline(yintercept = lower, color = "red", linetype = "dashed") +
 geom_hline(yintercept = upper, color = "red", linetype = "dashed") +
 ggtitle("Bland-Altman Plot") +
 xlab("Average Measurement") +
 ylab("Difference Between Measurements")



library(BlandAltmanLeh)
bland.altman.plot(ba_lung$Manual, ba_lung$Digital, conf.int = 0.95, graph.sys = "ggplot2")


library(ggplot2)
manual  <- lung_scores$manual
digital <- lung_scores$digital_tree_binned

diffs <- manual - digital
means <- (manual + digital) / 2

bias <- mean(diffs)
loa  <- bias + c(-1.96, 1.96) * sd(diffs)

ba_plot_lung <- ggplot(data.frame(means, diffs), aes(x = means, y = diffs)) +
  #geom_point(size = 3, alpha = 0.7, color = "#219ebc") +
  geom_jitter(size = 3, alpha = 0.7, color = "#FB8500") +
  geom_hline(yintercept = bias, color = "red", size = 1) +
  geom_hline(yintercept = loa, linetype = "dashed", color = "black") +
  annotate("text", x = 0.5, y = bias + 0.2,
           label = paste0("Bias = ", round(bias, 2))) +
  annotate("text", x = 0.5, y = loa[1] + 0.2,
           label = paste0("Lower LoA = ", round(loa[1], 2))) +
  annotate("text", x = 0.5, y = loa[2] - 0.2,
           label = paste0("Upper LoA = ", round(loa[2], 2))) +
  labs(title = "Bland–Altman Plot: Manual vs Digital (Binned)",
       x = "Mean of Methods",
       y = "Difference (Manual - Digital)") +
  theme_bw()

ba_plot_lung

```


```{r}
# 1. Bland–Altman analysis (gold standard)
# This is the primary method used in biomedical method comparison.
# 
# It tests interchangeability by evaluating:
# Mean bias (systematic difference)
# Limits of agreement (LoA)
# Heteroscedasticity (variance changes with magnitude)
# 
# Interpretation:
# Bias ≈ 0 → no systematic difference
# LoA narrow enough → methods interchangeable
# LoA wide → methods not interchangeable
# For your case, you will almost certainly see:
# A non-zero bias (manual is coarse, digital is continuous)
# Wide LoA (manual has higher measurement error)
# This supports your message:
# 
# “Good agreement, but not interchangeable.”

```


```{r}
two_panel <- ggarrange(ba_plot_lung,ba_plot_brain,
                       ncol = 2,
                       nrow = 1,
                       common.legend = FALSE)

two_panel

# Save the two panel plot
ggsave(paste0(output_dir, "lung_mq_vs_dq_&_brain_mq_vs_dq_bland_altman.pdf"), 
       two_panel, 
       width = 13, 
       height = 4.5, 
       dpi = 600)
```

# Lung
```{r}
# Bland–Altman analysis showed a minimal mean bias (–0.07), indicating no systematic difference between manual and digital (binned) scoring. However, the limits of agreement were wide (–0.79 to +0.66), meaning that individual measurements can differ by up to ~0.7–0.8 scoring units. This variability is substantial relative to the 0–4 scale and exceeds acceptable thresholds for interchangeability. Thus, although the two methods agree on average, they cannot be considered interchangeable. The digital method provides a more precise and consistent measurement, whereas the manual method introduces greater variability.
```

# Brain
```{r}
#The brain Bland–Altman plot shows minimal bias and relatively narrow limits of agreement (–0.38 to +0.38), indicating good concordance between manual and digital scoring. Although the two methods show closer agreement than in the lung dataset, the remaining variability suggests that they are not fully interchangeable. The digital method remains more precise and provides a more granular quantitative assessment.
```


```{r}
# The lung dataset (n=250) showed wider limits of agreement (–0.79 to +0.66) than the brain dataset (n=106; –0.38 to +0.38). This difference likely reflects both the larger sample size, which captures greater variability, and the inherently higher heterogeneity and scoring difficulty of lung pathology. In contrast, brain scores were more consistent across samples, resulting in narrower limits of agreement.
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

