---
title: "R CCC analysis"
output: html_notebook
---
This is the CCC analysis notebook, divided in two main parts *Brain* and *Lung*
It starts by Brain, loading the data, cleaning and merging the data, transforming the corresponding columns for the analysis and finding the best comparable scales among manual score [0-4] to digital scores (%), where in Brain we have a max of 20% while in Lung we have a max of 60%. Therefore, using a comparable scale to meet the assumptions for the CCC analysis, we will establish.
For *Brain* [0-4] corresponding to (-Inf,0,5,10,15,20,Inf)(%), while for *Lung* [0-4] we will establish to (-Inf,15,30,45,60,Inf)(%)
# Load libraries
```{r}
library(ggplot2)
library(ggpubr)
library(dplyr)
```


```{r}
gc()
rm(list = ls(all.names = TRUE))
output_dir <- "C:/Users/emari/OneDrive - Universitaet Bern/Nedim/2026/Results/CCC/"
if (!dir.exists(output_dir)){dir.create(output_dir,recursive = TRUE)}
```


## Reload brain data 
```{r}
sample_sheet = "C:/Users/emari/OneDrive - Universitaet Bern/Nedim/2026/data/All lungs and brain_cleaned_Dec2025_NK.xlsx"
sample_sheet_sheet_4_brain = readxl::read_excel(sample_sheet, col_names = TRUE, sheet = 4)
head(sample_sheet_sheet_4_brain)
colnames(sample_sheet_sheet_4_brain)


sample_sheet_sheet_6_brain = readxl::read_excel(sample_sheet, col_names = TRUE, sheet = 6)
head(sample_sheet_sheet_6_brain)
colnames(sample_sheet_sheet_6_brain)

sample_sheet_sheet_4_brain_coi <- sample_sheet_sheet_4_brain  |> dplyr::select(ID,`ID IVI`,`% Positive area`)
colnames(sample_sheet_sheet_4_brain_coi)
colnames(sample_sheet_sheet_6_brain)

# Need to merge colun % Positive Area from sheet 4 with Brain to plot sheet 6
sample_sheet_sheet_6_brain <- sample_sheet_sheet_6_brain |> dplyr::rename("ID"="ID...39")
sample_sheet_brain <- sample_sheet_sheet_6_brain |> dplyr::left_join(sample_sheet_sheet_4_brain_coi ,by = "ID")#c("ID...1"="ID"))

colnames(sample_sheet_brain)
head(sample_sheet_brain$`% Positive area.y`)


table(sample_sheet_brain$`Vaccinated (v), none (n)`)
table(sample_sheet_brain$`ID IVI.x`,sample_sheet_brain$`Euthanasia day (dpi)`)

sample_sheet_brain$Euthanasia_day_dpi = gsub(x = sample_sheet_brain$`Euthanasia day (dpi)`,
                                             pattern ="?p\\.v$",
                                             replacement = "")
sample_sheet_brain$Euthanasia_day_dpi <- as.numeric(sample_sheet_brain$Euthanasia_day_dpi)
  

message("using latest day of euthanasia day")
multima_day <- max(sample_sheet_brain$Euthanasia_day_dpi)

sample_sheet_brain <- sample_sheet_brain  |> dplyr::mutate(
    # Time group based on euthanasia day
    Time_Group = case_when(
      Euthanasia_day_dpi >= 0 & Euthanasia_day_dpi <= 4 ~ "Early infection (0-4 dpi)",
      Euthanasia_day_dpi >= 5 & Euthanasia_day_dpi <= 7 ~ "Peak of disease (5-7 dpi)",
      Euthanasia_day_dpi >= 8 & Euthanasia_day_dpi <= multima_day  ~ paste0("Resolution (8-",multima_day," dpi)"),
      TRUE ~ "Other"),
    # Vaccination status
    Vaccination_Status = if_else(`Vaccinated (v), none (n)` == "n", "n", "v") |> factor())#(n)`, levels = c("n", "v"))
  
sample_sheet_brain <- sample_sheet_brain |> mutate( Vaccination_Status = case_when( grepl(" p\\.v$", `Euthanasia day (dpi)`) ~ "v/nc", TRUE ~ as.character(Vaccination_Status) ),
                                                     Vaccination_Status = factor(Vaccination_Status, levels = c("n", "v", "v/nc")))

table(sample_sheet_brain$Vaccination_Status)
# Create combined grouping variable
sample_sheet_brain <- sample_sheet_brain  |> dplyr::mutate(Combined_Group = paste0(Time_Group, "-", Vaccination_Status))

sample_sheet_brain$sex <- as.factor(toupper(sample_sheet_brain$sex))


# Prepare data for comparison
comparison_data_brain <- sample_sheet_brain |> 
  dplyr::select(
    `Manual Score brain`,
    ID...1,
    sex,
    `qPCR swabs (RNA/ml)`,
    `Histopathological score brain`,
    #`Histopathological score brain...24`,
    `% Positive area.y`,
    Time_Group,
    Vaccination_Status,
    `qPCR Brain (RNA/ml)`,
    #qPCR brain log10,
    `Live Virus log10 brain`,
    `Live virus (PFU/mouse) brain`) |>
  drop_na(`% Positive area.y`)



# Rename for easier plotting
names(comparison_data_brain) <- c(
  "manual semiquantification brain",
  "ID",
  "sex",
  "qPCR_swabs",
  "Histopath_Brain",
  #"Histopath_Brain_24",
  "digital quantification brain",
  "Time_Group",
  "Vaccination_Status",
  "qPCR Brain (RNA/ml)",
  #qPCR brain log10,
  "Live Virus log10 brain",
  "Live virus (PFU/mouse) brain")


#change scores from 0-5 to 0-4 (leave the only animal with score 1 as it is and change the rest.)
comparison_data_brain$`manual semiquantification brain`

comparison_data_brain$`manual semiquantification brain` <- dplyr::recode(
  comparison_data_brain$`manual semiquantification brain`,
  `0` = 0,
  `2` = 1,
  `3` = 2,
  `4` = 3,
  `5` = 4
)

comparison_data_brain$`manual semiquantification brain`
#Treat v/nc as vaccinated --> converted
comparison_data_brain$Vaccination_Status
comparison_data_brain$Vaccination_Status <- if_else(comparison_data_brain$Vaccination_Status == "v/nc","v",comparison_data_brain$Vaccination_Status)

comparison_data_brain
```

```{r}
# Correct data from brain not log-ed.
message("Some values in the qPCR for swabs and Brain (RNA/ml) and Live virus, are converted to log10, therefore ok, while others have a vlaue that is huge and need to be converted.")
head(comparison_data_brain$qPCR_swabs)
head(comparison_data_brain$`qPCR Brain (RNA/ml)`)
head(comparison_data_brain$`Live virus (PFU/mouse) brain`)

```
```{r}
message("Tipycally values smaller than 100 are log10 converted...")
log_candidates <- c("qPCR_swabs","qPCR Brain (RNA/ml)", "Live virus (PFU/mouse) brain")
for (col in log_candidates) {
  comparison_data_brain[[col]] <- ifelse(
    comparison_data_brain[[col]] > 100,
    log10(comparison_data_brain[[col]]),
    comparison_data_brain[[col]])}
```

```{r}
head(comparison_data_brain$qPCR_swabs)
min(round(comparison_data_brain$qPCR_swabs[!is.na(comparison_data_brain$qPCR_swabs)],digits = 3))
max(round(comparison_data_brain$qPCR_swabs[!is.na(comparison_data_brain$qPCR_swabs)],digits = 3))
min(round(comparison_data_brain$qPCR_swabs,digits = 3))
head(comparison_data_brain$`qPCR Brain (RNA/ml)`)
min(round(comparison_data_brain$`qPCR Brain (RNA/ml)`[!is.na(comparison_data_brain$`qPCR Brain (RNA/ml)`)],digits = 5))
max(round(comparison_data_brain$`qPCR Brain (RNA/ml)`[!is.na(comparison_data_brain$`qPCR Brain (RNA/ml)`)],digits = 5))
#comparison_data$`qPCR Brain (RNA/ml)`[round(comparison_data$`qPCR Brain (RNA/ml)`[!is.na(comparison_data$`qPCR Brain (RNA/ml)`)],digits = 3) == 68.616]
#comparison_data$`qPCR Brain (RNA/ml)`[comparison_data$`qPCR Brain (RNA/ml)` == 68.616] found and converted to the real value from Nedim data.

head(comparison_data_brain$`Live virus (PFU/mouse) brain`)
```
```{r}
head(comparison_data_brain)
dim(comparison_data_brain)
comparison_data_brain
```


### Spearman
#### How variables change together? (meassured by correlation)

#### Brain manual correlation to digital
```{r}
brain_scores <- comparison_data_brain |>
  dplyr::select(
    manual = `manual semiquantification brain`,
    digital = `digital quantification brain`,
    vaccination_status = Vaccination_Status,
    time_group = Time_Group
  ) |>
  drop_na()

# Brain correlation in general manual to digital
cor_spearman_brain <- cor(brain_scores$manual, brain_scores$digital,method = "spearman", use = "complete.obs")
cor_spearman_brain
```

```{r}
ggplot(brain_scores, aes(x = manual, y = digital)) +
  geom_point(size = 3, alpha = 0.7, color = "darkblue") +
  geom_smooth(method = "lm", se = TRUE, color = "red", linewidth = 1) +
  ggpubr::stat_cor(
    method = "spearman",
    label.x.npc = "left",
    label.y.npc = "top",
    aes(label = paste0("Spearman = ", after_stat(r.label),
                       "\nP = ", after_stat(p.label)))) +
  labs(title = "Manual vs Digital Brain Scores",
    x = "Manual semiquantification (brain)",
    y = "Digital quantification (brain)") +
  theme_bw(base_size = 10)
```


```{r}
# Brain correlatios scores manual to digital by time group and vaccination status.
brain_scores_cor_brain <- list()
for (vs in brain_scores$vaccination_status){
  for (tg in brain_scores$time_group){
    df <- brain_scores |> dplyr::filter(vaccination_status == vs & time_group == tg)
    brain_scores_cor_brain[[paste0(vs)]][[paste0(tg)]] <- cor(df$manual,df$digital,method = "spearman", use = "complete.obs")}}
brain_scores_cor_brain
```

```{r}
ggplot(brain_scores, aes(x = manual, y = digital)) +
  geom_point(aes(color = time_group), size = 3, alpha = 0.7) +
  geom_smooth(aes(color = time_group), method = "lm", se = FALSE) +
  
  # Spearman correlation per facet
  ggpubr::stat_cor(
    method = "spearman",
    label.x.npc = "left",
    label.y.npc = "top",
    aes(label = paste0("S = ", after_stat(r.label),
                       "\nP = ", after_stat(p.label)))) +
  facet_grid(vaccination_status ~ time_group) +
  labs(title = "Manual vs Digital Brain Scores",
    x = "Manual semiquantification (brain)",
    y = "Digital quantification (brain)") +
  theme_bw(base_size = 10) +
  theme(legend.position = "none",
    strip.text = element_text(size = 10, face = "bold"))
```




# MI
#### Will tell you if the methods are related, but it cannot tell you if they are interchangeable. 
#### High MI just means the digital score is a good "predictor" of the manual score.

## Discrete
### Brain
```{r}
#install.packages("infotheo")
library(infotheo)

brain_disc <- discretize(brain_scores)
mi_value_brain_discrete <- mutinformation(brain_disc$manual, brain_disc$digital)
mi_value_brain_discrete
# need to run under before this..
# mi_value_brain_discrete_binned <- mutinformation(brain_disc$manual, brain_disc$digital_binned)
# mi_value_brain_discrete_binned
```


# CCC
### 1. Do the manual  and digital quantification agree, and can they be used interchangeably?  
### 2. Agreement measures how close the actual values are to one another


### 3.Convert the "%" scores from the digital to binary format.
```{r}
# Bin the digital scores to match the 0-5 scale
# Adjust 'breaks' based on your clinical expertise/standard
brain_scores$digital_binned <- cut(brain_scores$digital,
                                   #breaks = c(-Inf, 1, 5, 10, 15, 20, Inf),
                                   breaks = c(-Inf, 5, 10, 15, 20, Inf),
                                   labels = 0:4,
                                   include.lowest = TRUE)


# Ensure both are numeric or factors for the function
brain_scores$manual
brain_scores$digital_binned <- as.numeric(as.character(brain_scores$digital_binned))
max(brain_scores$digital_binned)
```


```{r}
#install.packages("DescTools")
library(DescTools)
#--> Results :  CCC estimate,accuracy,precision,95% CI

#ccc_result_brain <- CCC(brain_scores$manual, brain_scores$digital)
#for status_vaccinated in brain_scores
ccc_result_brain <- CCC(brain_scores$manual, brain_scores$digital_binned)
ccc_result_brain
round(ccc_result_brain$rho.c[,1],3)
ccc_result_brain$rho.c      # overall CCC
ccc_result_brain$s.shift    # scale shift
ccc_result_brain$l.shift    # location shift
ccc_result_brain$C.b        # accuracy
```

```{r}
write.table(comparison_data_brain,file = paste0(output_dir,"comparison_data_brain.csv"),sep = ",")
```


```{r}
brain_ccc <- ggplot(brain_scores, aes(x = manual, y = digital_binned)) +
  geom_jitter(size = 3, alpha = 0.7, color = "#3C1642") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  stat_cor(method = "spearman", label.x =0, label.y = 5,size=5) +
  annotate("text", 
           label = paste0("CCC = ", round(ccc_result_brain$rho.c[,1], 3)),
           x= 0.44,y=4.7,#hjust=-0.3,vjust=3,
           size=5) +
  xlim(0,4)+
  ylim(0,4)+
  labs(
    title = "Manual vs Digital Brain Scoring (Binnned)",
    x = "Manual Semiquantification",
    y = "Digital Quantification (% Positive Area)",
    subtitle = paste0("n=",brain_scores |> summarize(n=n())) #lung_scores |> count()
  ) +
  theme_bw()
brain_ccc
```


```{r}

```


```{r}
```


```{r}
```
## Reload data lung
```{r}
sample_sheet = "C:/Users/emari/OneDrive - Universitaet Bern/Nedim/2026/data/All lungs and brain_cleaned_Dec2025_NK.xlsx"
sample_sheet_lungs = readxl::read_excel(sample_sheet, col_names = TRUE, sheet = 2)
head(sample_sheet_lungs)
colnames(sample_sheet_lungs)
table(sample_sheet_lungs$`ID IVI`,sample_sheet_lungs$`Euthanasia day (dpi)`)

table(sample_sheet_lungs$`Vaccinated (v), none (n)`)
# sample_sheet_lungs$Euthanasia_day_dpi = gsub(x = sample_sheet_lungs$`Euthanasia day (dpi)`,
#                                              pattern ="?p\\.v$",
#                                              replacement = "")
sample_sheet_lungs$Euthanasia_day_dpi <- as.numeric(sample_sheet_lungs$`Euthanasia day (dpi)`)
mterm_val <- max(sample_sheet_lungs$Euthanasia_day_dpi)

#sample_sheet_lungs$`Vaccinated (v), none (n)`

message("Add to plot the number of samples such as: n=12")

sample_sheet_lungs <- sample_sheet_lungs  |> dplyr::mutate(
    # Time group based on euthanasia day
    Time_Group = case_when(
      Euthanasia_day_dpi >= 0 & Euthanasia_day_dpi <= 4 ~ "Early infection (0-4 dpi)",
      Euthanasia_day_dpi >= 5 & Euthanasia_day_dpi <= 7 ~ "Peak of disease (5-7 dpi)",
      Euthanasia_day_dpi >= 8 & Euthanasia_day_dpi <= mterm_val ~ paste0("Resolution (8-",mterm_val," dpi)"),
      TRUE ~ "Other"),
    # Vaccination status
    Vaccination_Status = if_else(`Vaccinated (v), none (n)` == "n", "n", "v") |> factor())

# Create combined grouping variable
sample_sheet_lungs <- sample_sheet_lungs  |> dplyr::mutate(
    Combined_Group = paste(Time_Group, "-", Vaccination_Status)
  )

table(sample_sheet_lungs$Vaccination_Status)

# Prepare data for comparison
comparison_data_lung <- sample_sheet_lungs |> 
  dplyr::select(
    `Manual Score (0 to 4)`,
    ID,
    `qPCR swabs (RNA/ml)`,
    `Histopathological score`,
    #`Histopathological score brain...24`,
    `% Positivity in total parenchyma`,
    Time_Group,
    Vaccination_Status,
    `qPCR Lung (RNA/ml)`,
    #qPCR brain log10,
    `Live virus (PFU/ml)`) #|>  drop_na(`% Positive area.y`)




# Rename for easier plotting
names(comparison_data_lung) <- c(
  "manual semiquantification lung",
  "ID",
  "qPCR_swabs",
  "Histopath_Lung",
  #"Histopath_Brain_24",
  "digital quantification lung",
  "Time_Group",
  "Vaccination_Status",
  "qPCR Lung (RNA/ml)",
  #qPCR brain log10,
  "Live virus (PFU/ml)")


head(comparison_data_lung)
dim(comparison_data_lung)
comparison_data_lung

```
# Spearman
### How variables change together? (meassured by correlation)
### Lung manual correlation to digital

```{r}
lung_scores <- comparison_data_lung |>
  dplyr::select(
    manual = `manual semiquantification lung`,
    digital = `digital quantification lung`,
    vaccination_status = Vaccination_Status,
    time_group = Time_Group)# |>drop_na()

cor_spearman_lung <- cor(lung_scores$manual, lung_scores$digital,
                    method = "spearman", use = "complete.obs")
cor_spearman_lung
```

```{r}
ggplot(lung_scores, aes(x = manual, y = digital)) +
  geom_point(size = 3, alpha = 0.7, color = "darkorange") +
  geom_smooth(method = "lm", se = TRUE, color = "red", linewidth = 1) +
  ggpubr::stat_cor(
    method = "spearman",
    label.x.npc = "left",
    label.y.npc = "top",
    aes(label = paste0("Spearman = ", after_stat(r.label),
                       "\nP = ", after_stat(p.label)))) +
  labs(title = "Manual vs Digital Lung Scores",
    x = "Manual semiquantification (lung)",
    y = "Digital quantification (lung)") +
  theme_bw(base_size = 10)
```

```{r}
ggplot(lung_scores, aes(x = manual, y = digital)) +
  geom_point(aes(color = time_group), size = 3, alpha = 0.7) +
  geom_smooth(aes(color = time_group), method = "lm", se = FALSE) +
  
  # Spearman correlation per facet
  ggpubr::stat_cor(
    method = "spearman",
    label.x.npc = "left",
    label.y.npc = "top",
    aes(label = paste0("S = ", after_stat(r.label),
                       "\nP = ", after_stat(p.label)))) +
  
  facet_grid(vaccination_status ~ time_group) +
  labs(title = "Manual vs Digital Lung Scores",
    x = "Manual semiquantification (lung)",
    y = "Digital quantification (lung)") +
  theme_bw(base_size = 10) +
  theme(legend.position = "none",
    strip.text = element_text(size = 10, face = "bold"))

```

```{r}
write.table(comparison_data_lung,file = paste0(output_dir,"comparison_data_lung.csv"),sep = ",")
```


```{r}
```



# MI
#### Will tell you if the methods are related, but it cannot tell you if they are interchangeable. 
#### High MI just means the digital score is a good "predictor" of the manual score.

## Discrete
### Lung
```{r}

lung_disc <- discretize(lung_scores)
mi_value_lung_discrete <- mutinformation(lung_disc$manual, lung_disc$digital)
mi_value_lung_discrete

```


```{r}
lung_scores$digital_binned <- cut(lung_scores$digital,
                                   breaks = c(-Inf, 10, 25, 40, 55, Inf),
                                   labels = 0:4,
                                   include.lowest = TRUE)


# Ensure both are numeric or factors for the function
lung_scores$manual
lung_scores$digital_binned <- as.numeric(as.character(lung_scores$digital_binned))
max(lung_scores$digital_binned)
```
# CCC
### 1. Do the manual  and digital quantification agree, and can they be used interchangeably?  
### 2. Agreement measures how close the actual values are to one another


### 3.Convert the "%" scores from the digital to binary format.
```{r}
#install.packages("DescTools")
library(DescTools)
#--> Results :  CCC estimate,accuracy,precision,95% CI

#ccc_result_lung <- CCC(lung_scores$manual, lung_scores$digital)
ccc_result_lung <- CCC(lung_scores$manual, lung_scores$digital_binned)
ccc_result_lung
round(ccc_result_lung$rho.c[,1],3)
ccc_result_lung$rho.c      # overall CCC
ccc_result_lung$s.shift    # scale shift
ccc_result_lung$l.shift    # location shift
ccc_result_lung$C.b        # accuracy
```
```{r}
library(ggplot2)
library(ggpubr)


lung_ccc <- ggplot(lung_scores, aes(x = manual, y = digital_binned)) +
  geom_jitter(size = 3, alpha = 0.7, color = "#FB8500") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  geom_smooth(method = "lm", se = TRUE, color = "black") +
  stat_cor(method = "spearman", label.x =0, label.y = 5,size=5) +
  annotate("text", 
           label = paste0("CCC = ", round(ccc_result_lung$rho.c[,1], 3)),
           x=0.44, y=4.7,size=5) +
  xlim(0,4)+
  ylim(0,4)+
  labs(
    title = "Manual vs Digital Lung Scoring (Binned)",
    x = "Manual Semiquantification",
    y = "Digital Quantification (% Positive Area)",
    subtitle = paste0("n=",lung_scores |> summarize(n=n())) #lung_scores |> count()
  ) +
  theme_bw()
lung_ccc
```


```{r,fig.width=11,fig.height=5}
# Combine into 2‑panel figure
two_panel <- ggarrange(brain_ccc, lung_ccc,
                       ncol = 2,
                       nrow = 1,
                       common.legend = FALSE)

two_panel

# Save the two panel plot
ggsave(paste0(output_dir, "brain_mq_vs_dq_&_lubg_mq_vs_dq_ccc.pdf"), 
       two_panel, 
       width = 11, 
       height = 5, 
       dpi = 600)
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

